Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-08-10 src/tex/ms.tex
Update content

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space from beyond-GR theory
            \item birefringence: enhance one and suppress another
            \item GR predicts there is no birefringence
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Assume GW are generated as GR described
        \item birefringence is considered during the propagation
        \item 
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item 
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item 
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
            \item Longer sampling
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

\section{Everything}
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence.

@@

For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
but also the distance to the binary and the strength of the birefringence.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

@@

Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog \textbf{[cite LIGO]}, GWTC-3,
was performed to constrain the plausible range of $\kappa$.
Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
interferometer and priors of the parameters provided. \citep{Ashton_2019}

@@

@@

On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $$\iota$$.
In this case, the posteriors can recover the GR distributions for the binary parameters,
and the most probable value of $$\kappa$$ is close to $$0$$, which means the birefringence is weak or absent, and GR can be recovered.

\begin{figure*}[h]
    \script{GW150914_corner.py}
    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the parameter estimations (PEs) done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure*}

@@

@@

>>>>

Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence.

@@

For previous studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
but also the distance to the binary and the strength of the birefringence.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is $0$.
This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

@@

To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{LIGO_2021}, GWTC-3, using Bilby,
a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
and priors for the parameters in question. \citep{Ashton_2019}

@@

\subsection{Result for GW150914}

\begin{figure*}[h]
    \script{GW150914_corner.py}
    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure*}

@@

On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $\iota$.
In this case, the posteriors can recover the GR distributions for the binary parameters,
and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.

\subsection{Result on GWTC-3}

@@

\subsection{Limitation}

\subsection{Future Study}

@@

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space
            \item birefringence: enhance one and suppress another
        \end{enumerate}
        \item Amplitude Ratio
        \begin{enumerate}
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item PE on GWTC-3 to constrain $\kappa$
        \item Assume GW are generated as GR described
        \item birefringence is considered during the propagation
        \item waveform modification
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item figure with frequency-dependent and frequency-independent
            \item posterior from $\cos\iota$ looks different
            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item figure of $\kappa$ of all events
            \item constraint on $\kappa$
            \item improvement of constraint on $\kappa$ compare to previous studies
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
        \end{enumerate}
    \end{enumerate}
\end{enumerate}
